\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2022-04-06 lin-vs-cen.tex
Boundary work

\newcommand{\defin}[1]{\emph{\color{brown}#1}}
\setlength{\parskip}{1ex}

@@

\begin{abstract}
    These are some notes on the relationships betwee linear colouring numbers and centred colouring numbers.
\end{abstract}

@@

Let $G$ be a simple undirected graph.  A \emph{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.  A vertex $v$ of $G$ is a \emph{centre} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.  A colouring $\varphi$ of $G$ is \emph{centred} if every connected subgraph of $G$ has a centre with respect to $\varphi$. A colouring $\varphi$ of $G$ is \emph{linear} if every simple path in $G$ has a centre with respect to $\varphi$.

The \emph{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.  We note that an easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$. The \emph{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.  Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.

The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet[Theorem~1]{kun.obrien.ea:polynomial} who showed the following result:

@@

For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  Such an edge is \defin{vertical} if $v=x$ and \defin{horizontal} if $w=y$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.

@@

Let $G$ be an $a\times b$ pseudogrid.  For an edge $vw$ of $G_{a\times b}$, we let $P_{vw}:=\overline{P}_{vw}-\{v,w\}$ denote the (possibly empty) subpath containing the internal vertices of $\overline{P}_{vw}$.  For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.  In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths is isomorphic to a $\le\! 1$-subdivision of $G_{a\times b}$.  We call $\mathcal{P}$ a \defin{grid-partition} of $G$.

@@

  \item We first delete rows and columns from $G$ so that each colour that appears in $G$ appears at least $C$ times, for some large constant $C$.  From this point on the goal is to construct $P$ so that it contains each colour at least twice.

@@

\todo[inline]{Consider constructing this set so that colours that appear on the boundary don't count.  This just makes the factor of $d$ jump to a $d+2$.}

\begin{lem}\label{only_frequent}
  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d|A|$.
\end{lem}

\begin{proof}
  The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.  If there exists no $A\subseteq \varphi(V(G))$ with $\varphi_{\mathcal{P}}^{-1}(A) < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma. Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|$ to eliminate all vertices with colours in $A$, as follows:
  \begin{compactitem}
    \item For each $v:=(i,j)\in V(G_{k})$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
    \item For each horizontal edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
    \item For each vertical edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
    \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
  \end{compactitem}
  At this point $|R|=|C|\le d|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.  In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.

  Now apply induction on $G_0$ to get a $k'\times k'$ pseudogrid with
  \[  k'\ge k_0-d|\varphi(V(G_0))|  \ge k-d|A|-d|\varphi(V(G_0))|\ge k - d|\varphi(V(G))|
  \]
  that satisfies the conditions of the lemma.
\end{proof}

\begin{lem}
  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid, let $X\subseteq V(G)$ denote the set of vertices in the first and last row and column of $G$, and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus X| \ge d|A|$.
\end{lem}

\begin{proof}

\end{proof}

@@

\begin{lem}\label{one_colour_per_object}
  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d\cdot|A|$. Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
  \begin{compactenum}[(i)]
    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)$ and edge set
  \[
    E(H) := \{ (\alpha,\mu)\in X\times Y: \alpha\in\varphi_\mathcal{P}(\mu) \}
  \]
  By \cref{d_hall}, $H$ contains a subgraph $M$ with $\deg_M(x)=d$ for each $\alpha\in X$ and $\deg_M(y)\le 1$ for each $\mu\in Y$.  For each edge $\alpha\mu$ in $M$, set $\phi(\mu):=\alpha$.  This defines $\phi(\mu)$ for any $\mu\in Y$ with $\deg_M(\mu)=1$.  For each $\mu\in Y$ with $\deg_M(\mu)=0$ set $\phi(\mu):=\perp$.
\end{proof}

@@

\begin{lem}\label{pick_up_two}
  Let $G$ be a $k\times k$ pseudogrid with $k\ge??$,  let $s,v,w,t$ be vertices of $G$ with $s$ in the first column of $G$, neither $v$ nor $w$ in the first or last column of $G$ and $t$ in the last column of $G$.  Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
\end{lem}

\begin{proof}
  Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_k)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.  
\end{proof}

\begin{lem}\label{pick_up_everything}
  Let $r\ge 12$ be divisible by $3$ and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.  Then $G$ contains a path that contains every vertex in $S$. 
\end{lem}

@@

  Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}$.  For any $\mu_1\in Q'$, 
  $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.  
  Since this is true for all $\mu_1\in Q$ we have
  \[  3(5r+1)^2 \ge |\VE(G_{3r/2}(\mu))| \ge \sum_{\mu_1\in Q} B_{r/2}(\mu_1)
  = 3|Q'|(r-2)^2 \enspace .
  \]
  which implies that $|Q'|\le (5r+1)^2/(r-2)^2$.  For $r > 2$, this bound is decreasing in $r$ and is less than $26$ for any $r> 11\sqrt{26} + 57$.  In particular, $|Q'|< 26$ for any $r\ge 114$.
\end{proof}

@@

\begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.  Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon$.  In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.  By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.  Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
\end{proof}

>>>>

\definecolor{brightmaroon}{rgb}{0.76, 0.13, 0.28}
\definecolor{linkblue}{rgb}{0, 0.337, 0.227}
\newcommand{\defin}[1]{\emph{\color{brightmaroon}#1}}
\setlength{\parskip}{1ex}

@@

\begin{abstract}
  We prove that linear colouring number of any $k\times k$ pseudogrid is $\Omega(k)$.  By an argument of Kun et al, this result gives a tighter upper bound on the treedepth of a graph as a function of its linear colouring number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear colouring number.
\end{abstract}

@@

Let $G$ be a simple undirected graph.  A \defin{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.  A vertex $v$ of $G$ is a \defin{centre} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.  A colouring $\varphi$ of $G$ is \defin{centred} if every connected subgraph of $G$ has a centre with respect to $\varphi$. A colouring $\varphi$ of $G$ is \defin{linear} if every simple path in $G$ has a centre with respect to $\varphi$.

The \defin{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.  We note that an easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$. The \defin{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.  Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.

The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet{kun.obrien.ea:polynomial} who showed the following result:

@@

For positive integers $a$ and $b$, the \defin{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  Such an edge is \defin{vertical} if $v=x$ and \defin{horizontal} if $w=y$.  For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.  The \defin{boundary} $\partial G_{a\times b}$ of $G_{a\times b}$ a the set of vertices in the first and last row and column and all edges incident to these vertices.

@@

Let $G$ be an $a\times b$ pseudogrid.  For an edge $vw$ of $G_{a\times b}$, we let $P_{vw}:=\overline{P}_{vw}-\{v,w\}$ denote the (possibly empty) subpath containing the internal vertices of $\overline{P}_{vw}$.  For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.  In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths is isomorphic to a $\le\! 1$-subdivision of $G_{a\times b}$.  We call $\mathcal{P}$ a \defin{grid-partition} of $G$.  The \defin{boundary} of $G$ is defined as $\partial G:=\bigcup_{\mu\in\partial G_{a\times b}} V(P_\mu)$.

@@

  \item We first delete rows and columns from $G$ so that each colour that appears in $G$ appears at least $d$ times in the interior of $G$, for some large constant $d$.  From this point on the goal is to construct $P$ so that it contains each colour at least twice.

@@

\begin{lem}\label{only_frequent}
  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/(d+2)$ colours.
  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus \partial G_k| \ge d|A|$.
\end{lem}

\begin{proof}
  The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.  If there exists no $A\subseteq \varphi(V(G))$ with $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G_k| < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.  Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le (d+2)|A|$ to eliminate all vertices with colours in $A$, as follows:
  \begin{compactitem}
    \item For each $v:=(i,j)\in V(G_{k})\setminus \partial G_k$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
    \item For each horizontal edge $vw\in E(G_{k\times k})\setminus \partial G_k$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
    \item For each vertical edge $vw\in E(G_{k\times k})\setminus\partial G_k$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
    \item We add the first and last row to $R$ and the first and last column to $C$.
    \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
  \end{compactitem}
  At this point $|R|=|C|\le (d+2)|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.  In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.

  Now apply induction on $G_0$ to get a $k'\times k'$ pseudogrid with
  \[
    k'\ge k_0-(d+2)|\varphi(V(G_0))|
      \ge k-(d+2)|A|-(d+2)|\varphi(V(G_0))|
      \ge k - (d+2)|\varphi(V(G))|
  \]
  that satisfies the conditions of the lemma.
\end{proof}

@@

\begin{lem}\label{one_colour_per_object}
  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus \partial G_k| \ge d\cdot|A|$. Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
  \begin{compactenum}[(i)]
    \item For each $\mu\in\partial G_{k}$, $\phi(\mu)=\perp$.
    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi_\mathcal{P}(\mu)$.
    \item For each $\alpha\in\varphi(V(G))$,
    $|\phi^{-1}(\alpha)|\ge d$.
    % $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)\setminus\partial G_k$ and edge set
  \[
    E(H) := \{ (\alpha,\mu)\in X\times Y: \alpha\in\varphi_\mathcal{P}(\mu) \}
  \]
  By \cref{d_hall}, $H$ contains a subgraph $M$ with $\deg_M(x)=d$ for each $\alpha\in X$ and $\deg_M(y)\le 1$ for each $\mu\in Y$.  For each edge $\alpha\mu$ in $M$, set $\phi(\mu):=\alpha$.  This defines $\phi(\mu)$ for any $\mu\in Y$ with $\deg_M(\mu)=1$.  For each $\mu\in Y$ with $\deg_M(\mu)=0$ set $\phi(\mu):=\perp$.
\end{proof}

@@

\begin{lem}\label{pick_up_two}
  Let $G$ be a $a\times a$ pseudogrid with $a\ge??$,  let $s,v,w,t$ be vertices of $G$ with $s$ in the first column of $G$, neither $v$ nor $w$ in the first or last column of $G$ and $t$ in the last column of $G$.  Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
\end{lem}

\begin{proof}
  Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_k)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
  todo[inline]{Finish this.}
\end{proof}

\begin{lem}\label{pick_up_everything}
  Let $r\ge ??$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(G)\setminus \partial G$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $R$} for each $v\in S$. Then $G$ contains a path that contains every vertex in $S$.
\end{lem}

\begin{proof}[Proof Sketch]
  We adjust things so that we get a set of disjoint boxes, each of which contains one or two points of $S$ in its interior.
  We do the scan-line thing depicted in \cref{outline}.
\end{proof}

@@

  Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu_1))\}$.  For any $\mu_1\in Q'$,
  $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.
  Since this is true for all $\mu_1\in Q$ we have
  \[  3(5r+1)^2 \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
  = 3|Q'|(r-2)^2 \enspace .
  \]
  which implies that $|Q'|\le (5r+1)^2/(r-2)^2$.  For $r > 2$, this bound is decreasing in $r$ and is less than $26$ for any $r> 11\sqrt{26} + 57$.  In particular, $|Q'|< 26$ for any $r\ge 114$.
\end{proof}

@@

\begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.  Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon-2$.  In particular, for some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.  By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.  Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
\end{proof}
